Development Of Parallel Matrix Operations

### 2011-02-15 paper/paper.tex
Add some more explinations of the results

@@

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   32609        &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   53546        & 0.609        & 0.609    & xxx        \\\hline
1000    & 2       &   15812        & 2.062        & 1.031    & -0.409     \\\hline
1000    & 4       &    8101        & 4.025        & 1.006    & -0.132     \\\hline
1000    & 8       &    4703        & 6.934        & 0.991    & -0.064     \\\hline
1000    & 8       &    4210        & 7.746        & 0.968    & -0.053     \\\hline
500     & seq     &    4797        &   xxx         & xxx       & xxx        \\\hline
500     & 1       &    2046        & 2.345         & 2.345     & xxx        \\\hline
500     & 2       &    2632        & 1.823         & 0.911      & 1.573      \\\hline
500     & 4       &    1411        & 3.400         & 0.850     & 0.586      \\\hline
500     & 7       &     953        & 5.034         & 0.719     & 0.377      \\\hline
500     & 8       &     789        & 6.080         & 0.760     & 0.298      \\\hline
100     & seq     &      71        &   xxx         & xxx       & xxx        \\\hline
100     & 1       &      62        & 1.145        & 1.145    & xxx        \\\hline
100     & 2       &      42        & 1.690        & 0.845    & 0.355     \\\hline
100     & 4       &      36        & 1.972        & 0.493    & 0.441     \\\hline
100     & 7       &      35        & 2.029        & 0.290    & 0.492     \\\hline
100     & 8       &      35        & 2.029        & 0.254    & 0.502     \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/run-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/eff-vs-proc.png}
\end{center}
The results report strong parallel speedup especially as our matrix 
size increases. The speedup for the $1000 X 1000$ matrix multiplication reports
near linear speedup, and a efficiency of $1.0$ exactly. The particularly poor performance
for small matrices can potentially be traced back to the slight over head
that comes with starting and executing a Parallel Java team. It is possible that the time taken during
the initial dispatch of the threads in the parallel implementation is greater
than any possible speed up achieved by using them for small matrices.

@@

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/running-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/eff-vs-proc.png}
\end{center}
Overall the results for the sequential Strassen and Parallel-Strassen
algorithms reported significantly faster run-times when compared to the
sequential and parallel regular algorithms. However the speedup
compared to the other matrix multiplication results is not quite as
impressive. Their are many factors which contribute to this result.
First of all, for matrices smaller than size of 128 by 128 both of the
Strassen implementations default back to the regular sequential matrix
multiplication. During testing it was discovered that the regular sequential
algorithm is faster for this special case. This is why the results show no
speedup for 100 by 100 matrix. It should also be noted that in our parallel
Strassen implementation the maximum number of steps which can parallelized is
seven. As shown in the results above, when moving to eight parallel threads
very little speedup is achieved. The Strassen algorithms are also specially
designed to operate on matrices which are of a size which is a power of two.
In order to take advantage of this for ever matrix, if the matrix is not a
power of two the matrix is padded with zero's and then unpadded after the
algorithm is complete. Since none of the matrices tested were of a size which
was a power of two, the results shown are examples of the worst possible
running time for the sequential and parallel Strassen algorithms.

@@

\section{Conclusions}

>>>>

As shown above, the parallel matrix-vector multiplication's report super
linear speedup's for large matrices and accompanying vectors. It should be
noted that larger N's were used in testing these algorithms vs the results
that follow. This was merely because this algorithm is far less memory and
computationally intensive. The run times with a smaller N were so small that
it was hard to discern if a speedup was occurring in some cases. It is possible,
as seen in the other results through out this paper, that the poor performance
for the smaller 1000 by 1000 matrix multiplied by a vector of length 1000 is
caused by the overhead associated with Parallel Java. The time it takes to
actually compute the result is shorter than the time it takes to initialize
and dispatch the threads from Parallel Java.

@@

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   32609        &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   53546        & 0.609        & 0.609    & xxx        \\\hline
1000    & 2       &   15812        & 2.062        & 1.031    & -0.409     \\\hline
1000    & 4       &    8101        & 4.025        & 1.006    & -0.132     \\\hline
1000    & 8       &    4703        & 6.934        & 0.991    & -0.064     \\\hline
1000    & 8       &    4210        & 7.746        & 0.968    & -0.053     \\\hline
500     & seq     &    4797        &   xxx         & xxx       & xxx        \\\hline
500     & 1       &    2046        & 2.345         & 2.345     & xxx        \\\hline
500     & 2       &    2632        & 1.823         & 0.911      & 1.573      \\\hline
500     & 4       &    1411        & 3.400         & 0.850     & 0.586      \\\hline
500     & 7       &     953        & 5.034         & 0.719     & 0.377      \\\hline
500     & 8       &     789        & 6.080         & 0.760     & 0.298      \\\hline
100     & seq     &      71        &   xxx         & xxx       & xxx        \\\hline
100     & 1       &      62        & 1.145        & 1.145    & xxx        \\\hline
100     & 2       &      42        & 1.690        & 0.845    & 0.355     \\\hline
100     & 4       &      36        & 1.972        & 0.493    & 0.441     \\\hline
100     & 7       &      35        & 2.029        & 0.290    & 0.492     \\\hline
100     & 8       &      35        & 2.029        & 0.254    & 0.502     \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/run-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel/eff-vs-proc.png}
\end{center}
The results report strong parallel speedup especially as our matrix 
size increases. The speedup for the 1000 by 1000 matrix multiplication reports
near linear speedup, and a efficiency of 1.0 exactly. The particularly poor performance
for small matrices can potentially be traced back to the slight over head
that comes with starting and executing a Parallel Java team. It is possible that the time taken during
the initial dispatch of the threads in the parallel implementation is greater
than any possible speed up achieved by using them for small matrices.

@@

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/running-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/eff-vs-proc.png}
\end{center}
Overall the results for the sequential Strassen and Parallel-Strassen
algorithms reported significantly faster run-times when compared to the
sequential and parallel regular algorithms. However the speedup
compared to the other matrix multiplication results is not quite as
impressive. Their are many factors which contribute to this result.
First of all, for matrices smaller than size of 128 by 128 both of the
Strassen implementations default back to the regular sequential matrix
multiplication. During testing it was discovered that the regular sequential
algorithm is faster for this special case. This is why the results show no
speedup for 100 by 100 matrix, and the first data point for the  100 by 100
matrix has such a extremely high efficiency. It should also be noted that in our parallel
Strassen implementation the maximum number of steps which can parallelized is
seven. As shown in the results above, when moving to eight parallel threads
very little speedup is achieved. The Strassen algorithms are also specially
designed to operate on matrices which are of a size which is a power of two.
In order to take advantage of this for ever matrix, if the matrix is not a
power of two the matrix is padded with zero's and then unpadded after the
algorithm is complete. Since none of the matrices tested were of a size which
was a power of two, the results shown are examples of the worst possible
running time for the sequential and parallel Strassen algorithms.

@@

These results are quite interesting in many ways. As observed in all of the
previous results, we see a increasing level of speedup as our problem size
increases. With the largest N performing the best, with a linear
speed up initially and slowly dropping in performance, until a drastic
crash in performance can be observed with running eight threads. This could
possibly be the final thread phenomenan discussed in class, where the
performance of the algorithm is reduced when the number of threads is equal
to the number of processors available. It would be interesting to continue
testing larger N's and seeing if the effect is more drastic the larger the N
as that appears to be what is happening with the relatively small data set.
\section{Conclusions}
